\documentclass{article}

\usepackage[margin=1in]{geometry}
\usepackage{longtable}

\title{Software Design Document (SDD)\\LA City Sidewalk Assessment Simulation}
\author{Elijah Rodriguez}
\date{\today}

\begin{document}

\maketitle
\tableofcontents
\newpage

\section{Revision History}

\begin{longtable}{|p{1.5in}|p{1in}|p{2.5in}|p{1in}|}
\hline
Name & Date & Reason for Change & Version \\ \hline
Elijah Rodriguez & ADD_DATE & Initial Snapshot 1 SDD draft & 1.0 \\ \hline
\end{longtable}

\section{Introduction}

\subsection{Purpose}

This Software Design Document (SDD) describes the architecture and internal system design for the LA City Sidewalk Assessment simulation project. The system is designed to support rover-based sidewalk inspection using LiDAR and point-cloud processing workflows.

\subsection{System Overview}

The system collects sidewalk scan data, processes LiDAR point clouds into orthoimages and elevation maps, uses a U-Net machine-learning model to identify sidewalk joints and cracks, calculates displacement measurements, and exports results for review.

\section{System Architecture}

The high-level system pipeline is:

Rover or iPhone LiDAR Scan $\rightarrow$ Raw Point Cloud $\rightarrow$ PointCloud2OrthoImage $\rightarrow$ Orthoimage and DEM/Elevation Image $\rightarrow$ U-Net Segmentation $\rightarrow$ Displacement Computation $\rightarrow$ CSV and Visualization Output $\rightarrow$ Reviewer Dashboard

\section{Module Breakdown}

\subsection{Data Acquisition Module}

Purpose: Collect sidewalk scan data using rover-mounted sensors and LiDAR-enabled devices.

Inputs:
\begin{itemize}
    \item LiDAR scans
    \item GPS data
    \item Sensor telemetry
\end{itemize}

Outputs:
\begin{itemize}
    \item Raw point-cloud files
    \item Scan metadata
\end{itemize}

\subsection{Point-Cloud Processing Module}

Purpose: Convert raw point-cloud data into orthoimages and DEM/elevation images.

Inputs:
\begin{itemize}
    \item LAS or PLY point-cloud files
\end{itemize}

Outputs:
\begin{itemize}
    \item Orthoimages
    \item DEM/elevation maps
\end{itemize}

\subsection{U-Net Segmentation Module}

Purpose: Detect sidewalk joints and cracks using image segmentation.

Inputs:
\begin{itemize}
    \item Orthoimages
\end{itemize}

Outputs:
\begin{itemize}
    \item Segmentation masks
\end{itemize}

\subsection{Displacement Computation Module}

Purpose: Calculate sidewalk displacement measurements from segmentation and elevation data.

Inputs:
\begin{itemize}
    \item DEM/elevation images
    \item Segmentation masks
\end{itemize}

Outputs:
\begin{itemize}
    \item CSV reports
    \item Visualization outputs
\end{itemize}

\subsection{Semi-Autonomous Navigation Module}

Purpose: Simulate GPS waypoint navigation and obstacle detection workflows.

Inputs:
\begin{itemize}
    \item GPS waypoints
    \item Obstacle detection events
\end{itemize}

Outputs:
\begin{itemize}
    \item Navigation commands
    \item Telemetry logs
\end{itemize}

\subsection{Storage and Reporting Module}

Purpose: Store processed data and generate reports for reviewers.

Inputs:
\begin{itemize}
    \item Processed outputs
    \item Logs and metadata
\end{itemize}

Outputs:
\begin{itemize}
    \item Reports
    \item Saved files
\end{itemize}

\section{Interfaces}

The system interfaces with rover hardware, LiDAR-enabled mobile devices, GPS systems, processing workstations, Docker services, and cloud or local storage systems.

\section{Design Constraints}

The system depends on LiDAR scan quality, point-cloud alignment accuracy, machine-learning model consistency, and compatible software dependencies such as TensorFlow, OpenCV, Laspy, and Open3D.

\section{Testing and Verification Strategy}

The system will be validated using TestRail reports, manual testing, segmentation validation, displacement verification, and workflow testing.

\section{Glossary}

\begin{itemize}
    \item ADA: Americans with Disabilities Act
    \item LiDAR: Light Detection and Ranging
    \item U-Net: Deep learning segmentation model
    \item DEM: Digital Elevation Model
\end{itemize}

\section{References}

\begin{itemize}
    \item LA City Sidewalk Assessment Project Materials
    \item TensorFlow Documentation
    \item OpenCV Documentation
    \item ROS Documentation
\end{itemize}

\end{document}